How do changes in ex vary between countries?

### 2022-03-10 draft_graphs.Rmd

# Now we create the change in ex variable as that's the response variable in the model
  f_e65_ENW <- f_e65_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

@@

@@

  #add labels
f_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")
```

>>>>

# Now we create the change in ex variable 
  f_e65_ENW <- f_e65_ENW %>% 
    arrange(year) %>% 
    mutate(delta_ex = ex - lag(ex)) %>% 
   ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
   geom_line() +
   geom_hline(yintercept = 0) + 
   stat_smooth(colour = "lightpink3")

@@

#   hmd_ex_selected_countries_with_synth %>% 
#       filter(sex == "male") %>% 
#       filter(code %in% c("GBRTENW", "GBR_SCO")) %>% 
#     group_by(code, x) %>% arrange(year) %>% 
#       mutate(delta_ex = ex - lag(ex)) %>%
#     ungroup %>% 
#     filter(year >= 1979) %>% 
#     ggplot(aes(x = year, y = delta_ex)) + geom_point() + 
#    geom_line() +
#    geom_hline(yintercept = 0) + 
#    stat_smooth(colour = "aquamarine4")+
# facet_grid(x ~ code)

  ex_ENW <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "GBRTENW") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "England and Wales", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")

@@

  #add labels
f_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_SCO + labs(y = "Annual change in e0", title = "Scotland", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_SCO + labs(y = "Annual change in e65", title = "Scotland", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_SCO <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "GBR_SCO") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Scotland", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_FRA + labs(y = "Annual change in e0", title = "France", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_FRA + labs(y = "Annual change in e65", title = "France", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_FRA <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "FRATNP") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "France", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ESP + labs(y = "Annual change in e0", title = "Spain", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ESP + labs(y = "Annual change in e65", title = "Spain", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_ESP <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "ESP") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Spain", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_ITA + labs(y = "Annual change in e0", title = "Italy", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_ITA + labs(y = "Annual change in e65", title = "Italy", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_ITA <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "ITA") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Italy", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTS + labs(y = "Annual change in e0", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTS + labs(y = "Annual change in e65", title = "Germany: Synthetic", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 #Grid
 ex_DEUTSYNTH <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "DEUTHSYNTH") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Germany: Synthetic", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTE + labs(y = "Annual change in e0", title = "Germany: East", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTE + labs(y = "Annual change in e65", title = "Germany: East", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_DEUTE <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "DEUTE") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Germany: East", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```

@@

  #add labels
f_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for females", caption = "Source: Human Mortality Database")
 m_e0_DEUTW + labs(y = "Annual change in e0", title = "Germany: West", subtitle = "Life expectancy changes at birth for males", caption = "Source: Human Mortality Database")
 f_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for females", caption = "Source: Human Mortality Database")
 m_e65_DEUTW + labs(y = "Annual change in e65", title = "Germany: West", subtitle = "Life expectancy changes at aged 65 for males", caption = "Source: Human Mortality Database")

 ex_DEUTW <- hmd_ex_selected_countries_with_synth %>%
      filter(sex %in% c("female", "male")) %>%
      filter(code == "DEUTW") %>%
    group_by(sex, x) %>% arrange(year) %>%
      mutate(delta_ex = ex - lag(ex)) %>%
    ungroup %>%
    filter(year >= 1979) %>%
    ggplot(aes(x = year, y = delta_ex)) + geom_point() +
   geom_line() +
   geom_hline(yintercept = 0) +
    stat_smooth(mapping = aes(colour = sex)) +
       facet_grid(x ~ sex)

  ex_ENW + labs(x = "Year", y = "Annual change in life expectancy", title = "Germany: West", subtitle = "Life expectancy changes at birth (0) and aged 65 (65) by sex", caption = "Source: Human Mortality Database")
```
